---
output: github_document
knit: (function(input, ...) { rmarkdown::render(input, output_file = "README.md", envir = globalenv()) })
---

### Contents

- [Team Standings]
- [Offensive and Defensive PPG]
- [Offensive and Defensive YPG]
- [Team Margins]
- [Point-Adjusted Margins]
- [Quarter-Based Scoring Trends]
- [Offensive and Defensive CPR]
- [Modeling]

```{r message = F, warning = F, include = F}
library(tidyverse)
library(tidymodels)
library(tvthemes)
library(janitor)
library(nflfastR)

theme_custom = theme_avatar() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 9, vjust = 2.5, face = "italic"),
        plot.caption = element_text(face = "italic"),
        panel.grid.major = element_line(linewidth = 0.5, colour = "#DFDAD1"),
        panel.grid.minor = element_line(linewidth = 0.5, colour = "#DFDAD1"))

theme_set(theme_custom)

write_csv(load_pbp(season = 2023), "data/season_pbp.csv") # doing this prevents data update message
season_pbp = read_csv("data/season_pbp.csv", show_col_types = F)

end_games = season_pbp |>
  filter(desc == "END GAME") |>
  select(game_id, date = game_date, week, away_team, away_score = total_away_score, home_score = total_home_score, home_team) |>
  mutate(win_team = case_when(home_score > away_score ~ home_team,
                              home_score < away_score ~ away_team,
                              home_score == away_score ~ "Tie"),
         win_score = case_when(home_score > away_score ~ home_score,
                               home_score < away_score ~ away_score,
                               home_score == away_score ~ home_score),
         lose_team = case_when(home_score > away_score ~ away_team,
                               home_score < away_score ~ home_team,
                               home_score == away_score ~ "Tie"),
         lose_score = case_when(home_score > away_score ~ away_score,
                                home_score < away_score ~ home_score,
                                home_score == away_score ~ away_score),
         game_margin = win_score - lose_score,
         total_points = win_score + lose_score)

all_teams = sort(unique(c(unique(end_games$home_team), unique(end_games$away_team))))

team_hex = teams_colors_logos |>
  select(team = team_abbr, team_color) |>
  filter(team %in% all_teams) |>
  pull(team_color)
```

___

### Team Standings

```{r echo = F}
get_n_wins = function(team) {
  return(end_games |> filter(win_team == team) |> nrow())
}

get_n_losses = function(team) {
  return(end_games |> filter(lose_team == team) |> nrow())
}

get_n_ties = function(team) {
  home_ties = end_games |> filter(win_team == "Tie" & home_team == team) |> nrow()
  away_ties = end_games |> filter(win_team == "Tie" & away_team == team) |> nrow()
  return(home_ties + away_ties)
}

data.frame(team = all_teams) |>
  mutate(wins = sapply(team, get_n_wins),
         losses = sapply(team, get_n_losses),
         ties = sapply(team, get_n_ties),
         win_pct = round(wins / (wins + losses), 3),
         record = paste0(wins, "-", losses, "-", ties),
         team2 = paste0(team, " (", record, ")")) |>
  ggplot(aes(reorder(team2, win_pct), win_pct)) +
  geom_col(aes(fill = team), show.legend = F) +
  scale_fill_manual(values = team_hex) +
  coord_flip() +
  labs(x = NULL, y = "Win Percentage", title = "Team Standings") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), labels = scales::percent)
```

___

### Offensive and Defensive PPG

```{r echo = F}
get_games_played = function(team) {
  x = c(end_games$home_team, end_games$away_team)
  return(length(x[x == team]))
}

get_pts_scored = function(team) {
  home_score = end_games |> filter(home_team == team) |> pull(home_score)
  away_score = end_games |> filter(away_team == team) |> pull(away_score)
  return(sum(c(home_score, away_score)))
}

get_pts_allowed = function(team) {
  home_score = end_games |> filter(home_team == team) |> pull(away_score)
  away_score = end_games |> filter(away_team == team) |> pull(home_score)
  return(sum(c(home_score, away_score)))
}

ppg_scatter = data.frame(team = all_teams) |>
  mutate(pts_scored = sapply(team, get_pts_scored),
         pts_allowed = sapply(team, get_pts_allowed),
         games_played = sapply(team, get_games_played),
         off_ppg = round(pts_scored / games_played, 3),
         def_ppg = round(pts_allowed / games_played, 3))

cor_lab = round(cor(ppg_scatter$off_ppg, ppg_scatter$def_ppg), 3)

ppg_scatter |>
  ggplot(aes(off_ppg, def_ppg)) +
  geom_point(aes(col = team), size = 5, shape = "square", show.legend = F) +
  ggrepel::geom_text_repel(aes(label = team), size = 3.5, box.padding = 0.5, segment.alpha = 0.5, max.overlaps = 32) +
  scale_color_manual(values = team_hex) +
  geom_line(stat = "smooth", formula = y ~ x, method = "lm", linetype = "dashed", alpha = 0.5) +
  labs(x = "Offensive PPG", y = "Defensive PPG",
       title = "Scatterplot of Offensive and Defensive PPG",
       subtitle = paste0("Correlation: ", cor_lab))
```

___

### Offensive and Defensive YPG

```{r echo = F}
ypg = season_pbp |>
  filter(!is.na(posteam)) |>
  group_by(team = posteam) |>
  summarise(off_yds = sum(yards_gained, na.rm = T)) |>
  mutate(games = sapply(team, get_games_played),
         off_ypg = round(off_yds / games, 3)) |>
  select(team, off_ypg) |>
  inner_join(season_pbp |>
  filter(!is.na(defteam)) |>
  group_by(team = defteam) |>
  summarise(def_yds = sum(yards_gained, na.rm = T)) |>
  mutate(games = sapply(team, get_games_played),
         def_ypg = round(def_yds / games, 3)) |>
  select(team, def_ypg), by = "team")

cor_lab = round(cor(ypg$off_ypg, ypg$def_ypg), 3)

ypg |>
  ggplot(aes(off_ypg, def_ypg)) +
  geom_point(aes(col = team), size = 5, shape = "square", show.legend = F) +
  scale_color_manual(values = team_hex) +
  ggrepel::geom_text_repel(aes(label = team), size = 3.5, box.padding = 0.5, segment.alpha = 0.5, max.overlaps = 32) +
  labs(x = "Offensive YPG", y = "Defensive YPG",
       title = "Scatterplot of Offensive and Defensive YPG",
       subtitle = paste0("Correlation: ", cor_lab)) +
  scale_x_continuous(breaks = seq(0, 500, by = 25)) +
  scale_y_continuous(breaks = seq(0, 500, by = 25)) +
  geom_line(stat = "smooth", formula = y ~ x, method = "lm", linetype = "dashed", alpha = 0.5)
```

___

### Team Margins

```{r echo = F}
margins = data.frame(team = all_teams) |>
  mutate(scored = sapply(team, get_pts_scored),
         allowed = sapply(team, get_pts_allowed),
         games = sapply(team, get_games_played)) |>
  transmute(team, margin = scored - allowed, avg_margin = round((scored - allowed) / games, 3),
            team2 = paste0(team, " (", margin, ")"))

margins |>
  mutate(pos_lab = ifelse(margin >= 0, team, ""),
         neg_lab = ifelse(margin < 0, team, "")) |>
  ggplot(aes(reorder(team2, margin), margin)) +
  geom_col(aes(fill = team2), show.legend = F) +
  geom_text(aes(label = pos_lab), size = 3, alpha = 0.25, hjust = -0.25) +
  geom_text(aes(label = neg_lab), size = 3, alpha = 0.25, hjust = 1.25) +
  scale_fill_manual(values = team_hex) +
  coord_flip(ylim = c(min(margins$margin) * 1.05, max(margins$margin) * 1.05)) +
  labs(x = NULL, y = "Margin", title = "Team Margins") +
  scale_y_continuous(breaks = seq(-1000, 1000, by = 5))
```

___

### Point-Adjusted Margins

```{r echo = F}
pam_df = data.frame(team = all_teams) |>
  mutate(scored = sapply(team, get_pts_scored),
         allowed = sapply(team, get_pts_allowed),
         margin = scored - allowed,
         total = scored + allowed,
         pam = round(margin / total, 4) * 10) |>
  select(team, pam) |>
  mutate(team2 = paste0(team, " (", pam, ")"))

pam_df |>
  ggplot(aes(reorder(team2, pam), pam)) +
  geom_col(aes(fill = team), show.legend = F) +
  scale_fill_manual(values = team_hex) +
  coord_flip() +
  labs(x = NULL, y = "PAM", title = "PAM: Point-Adjusted Margin",
       subtitle = "PAM = (Points Scored - Points Allowed) / Points Scored") +
  annotate("text", x = 1.5, y = max(pam_df$pam) * 0.8, label = "Data: NFL.com via {nflfastR}", size = 3, fontface = "italic") +
  scale_y_continuous(breaks = seq(-100, 100, by = 1))
```

___

### Quarter-Based Scoring Trends

```{r echo = F}
teams_qtrs = data.frame()

for (team in all_teams) {
  for (qtr in 1:4) {
    teams_qtrs = rbind(teams_qtrs, data.frame(team = team, qtr = qtr))
  }
}

teams_qtrs |>
  left_join(season_pbp |>
  filter(sp == 1) |>
  mutate(pts_scored = case_when(touchdown == 1 ~ 6,
                                field_goal_result == "made" ~ 3,
                                extra_point_result == "good" ~ 1,
                                two_point_conv_result == "success" ~ 2,
                                safety == 1 ~ 2,
                                T ~ 1000000),
         scoring_team = case_when(touchdown == 1 ~ td_team,
                                  field_goal_result == "made" ~ posteam,
                                  extra_point_result == "good" ~ posteam,
                                  two_point_conv_result == "success" ~ posteam,
                                  safety == 1 ~ defteam,
                                  T ~ "XXXXXXXXXXXXXXX"),
         new_qtr = ifelse(qtr == 5, 4, qtr)) |>
  group_by(team = scoring_team, qtr = new_qtr) |>
  summarise(pts = sum(pts_scored),
            .groups = "drop") |>
  group_by(team) |>
  mutate(pct = round(pts / sum(pts), 3)) |>
  ungroup(), by = c("team", "qtr")) |>
  mutate(pts = replace_na(pts, 0),
         pct = replace_na(pct, 0)) |>
  ggplot(aes(qtr, pct)) +
  geom_line(aes(col = team), linewidth = 2, show.legend = F) +
  scale_color_manual(values = team_hex) +
  facet_wrap(vars(team)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.5), labels = scales::percent) +
  labs(x = "Quarter", y = "Percentage of Points Scored",
       title = "When are teams scoring their points?",
       subtitle = "Overtime treated as 4Q")
```

___

### Offensive and Defensive CPR

```{r echo = F}
ppg_df = data.frame(team = all_teams) |>
  mutate(scored = sapply(team, get_pts_scored),
         allowed = sapply(team, get_pts_allowed),
         games = sapply(team, get_games_played)) |>
  transmute(team, ppg = round(scored / games, 3), papg = round(allowed / games, 3))

cpr_df = end_games |>
  select(home_team, home_score, away_team, away_score) |>
  inner_join(ppg_df, by = c("home_team" = "team")) |>
  rename(home_ppg = ppg, home_papg = papg) |>
  inner_join(ppg_df, by = c("away_team" = "team")) |>
  rename(away_ppg = ppg, away_papg = papg) |>
  mutate(home_exp = round((home_ppg + away_papg) / 2, 3),
         away_exp = round((away_ppg + home_papg) / 2, 3),
         home_off_cpr = home_score - home_exp,
         home_def_cpr = away_exp - away_score,
         away_off_cpr = away_score - away_exp,
         away_def_cpr = home_exp - home_score) |>
  select(home_team, home_score, away_team, away_score, home_off_cpr, home_def_cpr, away_off_cpr, away_def_cpr)

get_off_cpr = function(team) {
  home = cpr_df |> filter(home_team == team) |> pull(home_off_cpr)
  away = cpr_df |> filter(away_team == team) |> pull(away_off_cpr)
  return(round(mean(c(home, away)), 3))
}

get_def_cpr = function(team) {
  home = cpr_df |> filter(home_team == team) |> pull(home_def_cpr)
  away = cpr_df |> filter(away_team == team) |> pull(away_def_cpr)
  return(round(mean(c(home, away)), 3))
}

team_cpr = data.frame(team = all_teams) |>
  mutate(off_cpr = sapply(team, get_off_cpr),
         def_cpr = sapply(team, get_def_cpr),
         total_cpr = off_cpr + def_cpr)

team_cpr |>
  ggplot(aes(off_cpr, def_cpr)) +
  geom_point(aes(col = team), shape = "square", size = 5, show.legend = F) +
  scale_color_manual(values = team_hex) +
  ggrepel::geom_text_repel(aes(label = team), size = 3.5, box.padding = 0.5, segment.alpha = 0.5, max.overlaps = 32) +
  geom_hline(aes(yintercept = 0), linetype = "dashed", alpha = 0.25) +
  geom_vline(aes(xintercept = 0), linetype = "dashed", alpha = 0.25) +
  labs(x = "Offensive CPR", y = "Defensive CPR",
       title = "Scatterplot of Offensive and Defensive CPRs",
       subtitle = "Composite Performance Rating (CPR) is an Evaluation of Performance Over/Under Expected") +
  scale_x_continuous(breaks = seq(-10, 10, by = 1)) +
  scale_y_continuous(breaks = seq(-10, 10, by = 1))

cpr_ranks = team_cpr |>
  arrange(desc(total_cpr))
```

##### Five Best Total CPRs

1. `r cpr_ranks$team[1]`: `r cpr_ranks$total_cpr[1]`
2. `r cpr_ranks$team[2]`: `r cpr_ranks$total_cpr[2]`
3. `r cpr_ranks$team[3]`: `r cpr_ranks$total_cpr[3]`
4. `r cpr_ranks$team[4]`: `r cpr_ranks$total_cpr[4]`
5. `r cpr_ranks$team[5]`: `r cpr_ranks$total_cpr[5]`

##### Five Worst Total CPRs

1. `r cpr_ranks$team[32]`: `r cpr_ranks$total_cpr[32]`
2. `r cpr_ranks$team[31]`: `r cpr_ranks$total_cpr[31]`
3. `r cpr_ranks$team[30]`: `r cpr_ranks$total_cpr[30]`
4. `r cpr_ranks$team[29]`: `r cpr_ranks$total_cpr[29]`
5. `r cpr_ranks$team[28]`: `r cpr_ranks$total_cpr[28]`

___

### Modeling

```{r echo = F}
modeling_metrics = pam_df |>
  select(team, pam) |>
  inner_join(margins |>
  select(team, avg_margin), by = "team") |>
  inner_join(ppg_df, by = "team")

modeling_df = end_games |>
  transmute(home_team, away_team,
            home_win = ifelse(win_team == home_team, 1, 0)) |>
  inner_join(modeling_metrics, by = c("home_team" = "team")) |>
  rename(home_pam = pam,
         home_avg_margin = avg_margin,
         home_ppg = ppg,
         home_papg = papg) |>
  inner_join(modeling_metrics, by = c("away_team" = "team")) |>
  rename(away_pam = pam,
         away_avg_margin = avg_margin,
         away_ppg = ppg,
         away_papg = papg)

log_mod = glm(home_win ~ home_pam + home_avg_margin + home_ppg + home_papg +
                         away_pam + away_avg_margin + away_ppg + away_papg,
              data = modeling_df, family = "binomial")

mod_res = modeling_df |>
  mutate(pred_prob = round(predict(log_mod, newdata = modeling_df, type = "response"), 3),
         pred_home_win = ifelse(pred_prob >= 0.5, 1, 0)) |>
  count(home_win == pred_home_win) |>
  pull(n)

acc_pct = paste0(round(mod_res[2] / sum(mod_res), 4) * 100, "%")
```

- First draft basic logistic regression accuracy: `r acc_pct`








