---
output: github_document
knit: (function(input, ...) { rmarkdown::render(input, output_file = "README.md", envir = globalenv()) })
---

##### *Formatting is off right now. Will be adjusting soon :)*

### Contents

- [Team Standings]
- [Offensive and Defensive PPG]
- [Offensive and Defensive YPG]
- [Team Margins]
- [Point-Adjusted Margins]
- [Quarter-Based Scoring Trends]
- [Offensive and Defensive CPR]
- [Weekly QB CER]
- [QB Air Yards v YAC]
- [Modeling]

```{r message = F, warning = F, include = F}
library(tidyverse)
library(tidymodels)
library(tvthemes)
library(janitor)
library(nflfastR)

theme_custom = theme_avatar() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 9, vjust = 2.5, face = "italic"),
        plot.caption = element_text(face = "italic"),
        panel.grid.major = element_line(linewidth = 0.5, colour = "#DFDAD1"),
        panel.grid.minor = element_line(linewidth = 0.5, colour = "#DFDAD1"))

theme_set(theme_custom)

write_csv(load_pbp(season = 2023), "data/season_pbp.csv") # doing this prevents data update message
season_pbp = read_csv("data/season_pbp.csv", show_col_types = F)

end_games = season_pbp |>
  filter(desc == "END GAME") |>
  select(game_id, date = game_date, week, away_team, away_score = total_away_score, home_score = total_home_score, home_team) |>
  mutate(win_team = case_when(home_score > away_score ~ home_team,
                              home_score < away_score ~ away_team,
                              home_score == away_score ~ "Tie"),
         win_score = case_when(home_score > away_score ~ home_score,
                               home_score < away_score ~ away_score,
                               home_score == away_score ~ home_score),
         lose_team = case_when(home_score > away_score ~ away_team,
                               home_score < away_score ~ home_team,
                               home_score == away_score ~ "Tie"),
         lose_score = case_when(home_score > away_score ~ away_score,
                                home_score < away_score ~ home_score,
                                home_score == away_score ~ away_score),
         game_margin = win_score - lose_score,
         total_points = win_score + lose_score)

all_teams = sort(unique(c(unique(end_games$home_team), unique(end_games$away_team))))

team_hex = teams_colors_logos |>
  select(team = team_abbr, team_color) |>
  filter(team %in% all_teams) |>
  pull(team_color)
```

___

### Team Standings

```{r echo = F}
get_n_wins = function(team) {
  return(end_games |> filter(win_team == team) |> nrow())
}

get_n_losses = function(team) {
  return(end_games |> filter(lose_team == team) |> nrow())
}

get_n_ties = function(team) {
  home_ties = end_games |> filter(win_team == "Tie" & home_team == team) |> nrow()
  away_ties = end_games |> filter(win_team == "Tie" & away_team == team) |> nrow()
  return(home_ties + away_ties)
}

data.frame(team = all_teams) |>
  mutate(wins = sapply(team, get_n_wins),
         losses = sapply(team, get_n_losses),
         ties = sapply(team, get_n_ties),
         win_pct = round(wins / (wins + losses), 3),
         record = paste0(wins, "-", losses, "-", ties),
         team2 = paste0(team, " (", record, ")")) |>
  ggplot(aes(reorder(team2, win_pct), win_pct)) +
  geom_col(aes(fill = team), show.legend = F) +
  scale_fill_manual(values = team_hex) +
  coord_flip() +
  labs(x = NULL, y = "Win Percentage", title = "Team Standings") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), labels = scales::percent)
```

___

### Offensive and Defensive PPG

```{r echo = F}
get_games_played = function(team) {
  x = c(end_games$home_team, end_games$away_team)
  return(length(x[x == team]))
}

get_pts_scored = function(team) {
  home_score = end_games |> filter(home_team == team) |> pull(home_score)
  away_score = end_games |> filter(away_team == team) |> pull(away_score)
  return(sum(c(home_score, away_score)))
}

get_pts_allowed = function(team) {
  home_score = end_games |> filter(home_team == team) |> pull(away_score)
  away_score = end_games |> filter(away_team == team) |> pull(home_score)
  return(sum(c(home_score, away_score)))
}

ppg_scatter = data.frame(team = all_teams) |>
  mutate(pts_scored = sapply(team, get_pts_scored),
         pts_allowed = sapply(team, get_pts_allowed),
         games_played = sapply(team, get_games_played),
         off_ppg = round(pts_scored / games_played, 3),
         def_ppg = round(pts_allowed / games_played, 3))

ppg_scatter |>
  ggplot(aes(off_ppg, def_ppg)) +
  geom_point(aes(col = team), size = 5, shape = "square", show.legend = F) +
  ggrepel::geom_text_repel(aes(label = team), size = 3.5, box.padding = 0.5, segment.alpha = 0.5, max.overlaps = 32) +
  scale_color_manual(values = team_hex) +
  geom_hline(aes(yintercept = mean(def_ppg)), linetype = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = mean(off_ppg)), linetype = "dashed", alpha = 0.5) +
  labs(x = "Offensive PPG", y = "Defensive PPG",
       title = "Scatterplot of Offensive and Defensive PPG") +
  scale_x_continuous(breaks = seq(0, 50, by = 2.5)) +
  scale_y_continuous(breaks = seq(0, 50, by = 2.5))
```

___

### Offensive and Defensive YPG

```{r echo = F}
ypg = season_pbp |>
  filter(!is.na(posteam)) |>
  group_by(team = posteam) |>
  summarise(off_yds = sum(yards_gained, na.rm = T)) |>
  mutate(games = sapply(team, get_games_played),
         off_ypg = round(off_yds / games, 3)) |>
  select(team, off_ypg) |>
  inner_join(season_pbp |>
  filter(!is.na(defteam)) |>
  group_by(team = defteam) |>
  summarise(def_yds = sum(yards_gained, na.rm = T)) |>
  mutate(games = sapply(team, get_games_played),
         def_ypg = round(def_yds / games, 3)) |>
  select(team, def_ypg), by = "team")

ypg |>
  ggplot(aes(off_ypg, def_ypg)) +
  geom_point(aes(col = team), size = 5, shape = "square", show.legend = F) +
  scale_color_manual(values = team_hex) +
  ggrepel::geom_text_repel(aes(label = team), size = 3.5, box.padding = 0.5, segment.alpha = 0.5, max.overlaps = 32) +
  labs(x = "Offensive YPG", y = "Defensive YPG",
       title = "Scatterplot of Offensive and Defensive YPG") +
  scale_x_continuous(breaks = seq(0, 500, by = 25)) +
  scale_y_continuous(breaks = seq(0, 500, by = 25)) +
  geom_hline(aes(yintercept = mean(def_ypg)), linetype = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = mean(off_ypg)), linetype = "dashed", alpha = 0.5)
```

___

### Team Margins

```{r echo = F}
margins = data.frame(team = all_teams) |>
  mutate(scored = sapply(team, get_pts_scored),
         allowed = sapply(team, get_pts_allowed),
         games = sapply(team, get_games_played)) |>
  transmute(team, margin = scored - allowed, avg_margin = round((scored - allowed) / games, 3),
            team2 = paste0(team, " (", margin, ")"))

margins |>
  mutate(pos_lab = ifelse(margin >= 0, team, ""),
         neg_lab = ifelse(margin < 0, team, "")) |>
  ggplot(aes(reorder(team2, margin), margin)) +
  geom_col(aes(fill = team2), show.legend = F) +
  geom_text(aes(label = pos_lab), size = 3, alpha = 0.25, hjust = -0.25) +
  geom_text(aes(label = neg_lab), size = 3, alpha = 0.25, hjust = 1.25) +
  scale_fill_manual(values = team_hex) +
  coord_flip(ylim = c(min(margins$margin) * 1.05, max(margins$margin) * 1.05)) +
  labs(x = NULL, y = "Margin", title = "Team Margins") +
  scale_y_continuous(breaks = seq(-1000, 1000, by = 10))
```

___

### Point-Adjusted Margins

```{r echo = F}
pam_df = data.frame(team = all_teams) |>
  mutate(scored = sapply(team, get_pts_scored),
         allowed = sapply(team, get_pts_allowed),
         margin = scored - allowed,
         total = scored + allowed,
         pam = round(margin / total, 4) * 10) |>
  select(team, pam) |>
  mutate(team2 = paste0(team, " (", pam, ")"))

pam_df |>
  ggplot(aes(reorder(team2, pam), pam)) +
  geom_col(aes(fill = team), show.legend = F) +
  scale_fill_manual(values = team_hex) +
  coord_flip() +
  labs(x = NULL, y = "PAM", title = "PAM: Point-Adjusted Margin",
       subtitle = "PAM = (Points Scored - Points Allowed) / Points Scored") +
  annotate("text", x = 1.5, y = max(pam_df$pam) * 0.8, label = "Data: NFL.com via {nflfastR}", size = 3, fontface = "italic") +
  scale_y_continuous(breaks = seq(-100, 100, by = 1))
```

___

### Quarter-Based Scoring Trends

```{r echo = F}
teams_qtrs = data.frame()

for (team in all_teams) {
  for (qtr in 1:4) {
    teams_qtrs = rbind(teams_qtrs, data.frame(team = team, qtr = qtr))
  }
}

teams_qtrs |>
  left_join(season_pbp |>
  filter(sp == 1) |>
  mutate(pts_scored = case_when(touchdown == 1 ~ 6,
                                field_goal_result == "made" ~ 3,
                                extra_point_result == "good" ~ 1,
                                two_point_conv_result == "success" ~ 2,
                                safety == 1 ~ 2,
                                T ~ 1000000),
         scoring_team = case_when(touchdown == 1 ~ td_team,
                                  field_goal_result == "made" ~ posteam,
                                  extra_point_result == "good" ~ posteam,
                                  two_point_conv_result == "success" ~ posteam,
                                  safety == 1 ~ defteam,
                                  T ~ "XXXXXXXXXXXXXXX"),
         new_qtr = ifelse(qtr == 5, 4, qtr)) |>
  group_by(team = scoring_team, qtr = new_qtr) |>
  summarise(pts = sum(pts_scored),
            .groups = "drop") |>
  group_by(team) |>
  mutate(pct = round(pts / sum(pts), 3)) |>
  ungroup(), by = c("team", "qtr")) |>
  mutate(pts = replace_na(pts, 0),
         pct = replace_na(pct, 0)) |>
  ggplot(aes(qtr, pct)) +
  geom_line(aes(col = team), linewidth = 2, show.legend = F) +
  scale_color_manual(values = team_hex) +
  facet_wrap(vars(team)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.5), labels = scales::percent) +
  labs(x = "Quarter", y = "Percentage of Points Scored",
       title = "When are teams scoring their points?",
       subtitle = "Overtime treated as 4Q")
```

___

### Offensive and Defensive CPR

```{r echo = F}
ppg_df = data.frame(team = all_teams) |>
  mutate(scored = sapply(team, get_pts_scored),
         allowed = sapply(team, get_pts_allowed),
         games = sapply(team, get_games_played)) |>
  transmute(team, ppg = round(scored / games, 3), papg = round(allowed / games, 3))

cpr_df = end_games |>
  select(home_team, home_score, away_team, away_score) |>
  inner_join(ppg_df, by = c("home_team" = "team")) |>
  rename(home_ppg = ppg, home_papg = papg) |>
  inner_join(ppg_df, by = c("away_team" = "team")) |>
  rename(away_ppg = ppg, away_papg = papg) |>
  mutate(home_exp = round((home_ppg + away_papg) / 2, 3),
         away_exp = round((away_ppg + home_papg) / 2, 3),
         home_off_cpr = home_score - home_exp,
         home_def_cpr = away_exp - away_score,
         away_off_cpr = away_score - away_exp,
         away_def_cpr = home_exp - home_score) |>
  select(home_team, home_score, away_team, away_score, home_off_cpr, home_def_cpr, away_off_cpr, away_def_cpr)

get_off_cpr = function(team) {
  home = cpr_df |> filter(home_team == team) |> pull(home_off_cpr)
  away = cpr_df |> filter(away_team == team) |> pull(away_off_cpr)
  return(round(mean(c(home, away)), 3))
}

get_def_cpr = function(team) {
  home = cpr_df |> filter(home_team == team) |> pull(home_def_cpr)
  away = cpr_df |> filter(away_team == team) |> pull(away_def_cpr)
  return(round(mean(c(home, away)), 3))
}

team_cpr = data.frame(team = all_teams) |>
  mutate(off_cpr = sapply(team, get_off_cpr),
         def_cpr = sapply(team, get_def_cpr),
         total_cpr = off_cpr + def_cpr)

team_cpr |>
  ggplot(aes(off_cpr, def_cpr)) +
  geom_point(aes(col = team), shape = "square", size = 5, show.legend = F) +
  scale_color_manual(values = team_hex) +
  ggrepel::geom_text_repel(aes(label = team), size = 3.5, box.padding = 0.5, segment.alpha = 0.5, max.overlaps = 32) +
  geom_hline(aes(yintercept = 0), linetype = "dashed", alpha = 0.25) +
  geom_vline(aes(xintercept = 0), linetype = "dashed", alpha = 0.25) +
  labs(x = "Offensive CPR", y = "Defensive CPR",
       title = "Scatterplot of Offensive and Defensive CPRs",
       subtitle = "Composite Performance Rating (CPR) is an Evaluation of Performance Over/Under Expected") +
  scale_x_continuous(breaks = seq(-10, 10, by = 1)) +
  scale_y_continuous(breaks = seq(-10, 10, by = 1))

cpr_ranks = team_cpr |>
  arrange(desc(total_cpr))
```

##### Five Best Total CPRs

1. `r cpr_ranks$team[1]`: `r cpr_ranks$total_cpr[1]`
2. `r cpr_ranks$team[2]`: `r cpr_ranks$total_cpr[2]`
3. `r cpr_ranks$team[3]`: `r cpr_ranks$total_cpr[3]`
4. `r cpr_ranks$team[4]`: `r cpr_ranks$total_cpr[4]`
5. `r cpr_ranks$team[5]`: `r cpr_ranks$total_cpr[5]`

##### Five Worst Total CPRs

1. `r cpr_ranks$team[32]`: `r cpr_ranks$total_cpr[32]`
2. `r cpr_ranks$team[31]`: `r cpr_ranks$total_cpr[31]`
3. `r cpr_ranks$team[30]`: `r cpr_ranks$total_cpr[30]`
4. `r cpr_ranks$team[29]`: `r cpr_ranks$total_cpr[29]`
5. `r cpr_ranks$team[28]`: `r cpr_ranks$total_cpr[28]`

___

### Weekly QB CER

```{r echo = F}
current_week = 5

season_pbp |>
  filter(!is.na(air_yards) & week == current_week - 1) |>
  mutate(is_complete = ifelse(is.na(yards_after_catch), 0, 1)) |>
  select(passer, air_yards, is_complete) |>
  group_by(passer) |>
  summarise(n = n(),
            total_air = sum(air_yards),
            avg_air = mean(air_yards),
            comp_rate = mean(is_complete)) |>
  filter(n >= 10) |>
  transmute(passer, cer = round(avg_air / (1 - comp_rate), 3),
            cer = round(log(cer), 3),
            passer = paste0(passer, " (", cer, ")")) |>
  ggplot(aes(reorder(passer, cer), cer)) +
  geom_col(aes(fill = cer), show.legend = F) +
  coord_flip() +
  scale_fill_gradient(low = "indianred3", high = "springgreen4") +
  labs(x = NULL, y = "QB CER",
       title = "Past Week CER Rankings",
       subtitle = "CER (Chad Efficiency Rating) = ln(Avg. Air Yards / (1 - Completion Rate))") +
  scale_y_continuous(breaks = seq(0, 10, by = 0.5))
```

___

### QB Air Yards v YAC

```{r echo = F}
qb_air = season_pbp |>
  filter(!is.na(air_yards) & !is.na(yards_after_catch)) |>
  mutate(passer = paste0(substr(passer, 1, 2), " ", substr(passer, 3, nchar(passer)))) |>
  select(passer, air_yards, yards_after_catch) |>
  group_by(passer) |>
  summarise(n = n(),
            total_air = sum(air_yards),
            total_yac = sum(yards_after_catch)) |>
  slice_max(n, n = 32, with_ties = F) |>
  mutate(total_yds = total_air + total_yac,
         pct_air = round(total_air / total_yds, 3),
         pct_yac = round(total_yac / total_yds, 3)) |>
  select(passer, pct_air, pct_yac)

season_pbp |>
  filter(!is.na(air_yards) & !is.na(yards_after_catch)) |>
  mutate(passer = paste0(substr(passer, 1, 2), " ", substr(passer, 3, nchar(passer)))) |>
  select(passer, air_yards, yards_after_catch) |>
  group_by(passer) |>
  summarise(n = n(),
            total_air = sum(air_yards),
            total_yac = sum(yards_after_catch)) |>
  slice_max(n, n = 32, with_ties = F) |>
  mutate(total_yds = total_air + total_yac,
         pct_air = round(total_air / total_yds, 3),
         pct_yac = round(total_yac / total_yds, 3)) |>
  select(passer, pct_air, pct_yac) |>
  pivot_longer(!passer, names_to = "air_yac", values_to = "value") |>
  inner_join(qb_air, by = "passer") |>
  mutate(air_yac = ifelse(air_yac == "pct_yac", "Yards After Catch", "Air Yards"),
         air_yac = factor(air_yac, levels = c("Yards After Catch", "Air Yards")),
         passer = paste0(passer, " (", pct_yac * 100, "% YAC)")) |>
  ggplot(aes(reorder(passer, pct_air), value)) +
  geom_col(aes(fill = air_yac), width = 0.6) +
  scale_fill_manual(values = c("#D49898", "#85A67E")) +
  coord_flip() +
  labs(x = NULL, y = "Percent of Total Yards", fill = NULL,
       title = "Which QBs are YAC merchants?") +
  theme(legend.position = "right") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), labels = scales::percent)
```

___

### Modeling

```{r echo = F, message = F, warning = F}
modeling_metrics = pam_df |>
  select(team, pam) |>
  inner_join(margins |>
  select(team, avg_margin), by = "team") |>
  inner_join(ppg_df, by = "team")

modeling_df = end_games |>
  transmute(home_team, away_team,
            home_win = ifelse(win_team == home_team, 1, 0)) |>
  inner_join(modeling_metrics, by = c("home_team" = "team")) |>
  rename(home_pam = pam,
         home_avg_margin = avg_margin,
         home_ppg = ppg,
         home_papg = papg) |>
  inner_join(modeling_metrics, by = c("away_team" = "team")) |>
  rename(away_pam = pam,
         away_avg_margin = avg_margin,
         away_ppg = ppg,
         away_papg = papg)

log_mod = glm(home_win ~ home_pam + home_avg_margin + home_ppg + home_papg +
                         away_pam + away_avg_margin + away_ppg + away_papg,
              data = modeling_df, family = "binomial")

mod_res = modeling_df |>
  mutate(pred_prob = round(predict(log_mod, newdata = modeling_df, type = "response"), 3),
         pred_home_win = ifelse(pred_prob >= 0.5, 1, 0)) |>
  count(home_win == pred_home_win) |>
  pull(n)

acc_pct = paste0(round(mod_res[2] / sum(mod_res), 4) * 100, "%")
```

First draft basic logistic regression accuracy: `r acc_pct`

##### *This Week's Predictions*

```{r echo = F, message = F, warning = F}
predict_game_winner = function(home_team, away_team) {
  matchup_stats = data.frame(home_team = home_team, away_team = away_team) |>
    inner_join(pam_df, by = c("home_team" = "team")) |>
    select(home_team, away_team, home_pam = pam) |>
    inner_join(margins, by = c("home_team" = "team")) |>
    select(home_team, away_team, home_pam, home_avg_margin = avg_margin) |>
    inner_join(ppg_df, by = c("home_team" = "team")) |>
    rename(home_ppg = ppg, home_papg = papg) |>
    inner_join(pam_df, by = c("away_team" = "team")) |>
    select(home_team:home_papg, away_pam = pam) |>
    inner_join(margins, by = c("away_team" = "team")) |>
    select(home_team:away_pam, away_avg_margin = avg_margin) |>
    inner_join(ppg_df, by = c("away_team" = "team")) |>
    rename(away_ppg = ppg, away_papg = papg)
  
  home_win_prob = predict(log_mod, matchup_stats, type = "response")
  
  if (home_win_prob >= 0.5) {
    return(home_team)
  } else {
    return(away_team)
  }
}

predict_winner_prob = function(home_team, away_team) {
  matchup_stats = data.frame(home_team = home_team, away_team = away_team) |>
    inner_join(pam_df, by = c("home_team" = "team")) |>
    select(home_team, away_team, home_pam = pam) |>
    inner_join(margins, by = c("home_team" = "team")) |>
    select(home_team, away_team, home_pam, home_avg_margin = avg_margin) |>
    inner_join(ppg_df, by = c("home_team" = "team")) |>
    rename(home_ppg = ppg, home_papg = papg) |>
    inner_join(pam_df, by = c("away_team" = "team")) |>
    select(home_team:home_papg, away_pam = pam) |>
    inner_join(margins, by = c("away_team" = "team")) |>
    select(home_team:away_pam, away_avg_margin = avg_margin) |>
    inner_join(ppg_df, by = c("away_team" = "team")) |>
    rename(away_ppg = ppg, away_papg = papg)
  
  home_win_prob = predict(log_mod, matchup_stats, type = "response")
  
  if (home_win_prob >= 0.5) {
    return(round(home_win_prob, 3))
  } else {
    return(round(1 - home_win_prob, 3))
  }
}

this_week = nflfastR::fast_scraper_schedules(season = 2023) |>
  filter(week == 5) |>
  select(home_team, away_team) |>
  rowwise() |>
  mutate(pred_winner = predict_game_winner(home_team, away_team),
         pred_loser = ifelse(pred_winner == home_team, away_team, home_team),
         winner_prob = predict_winner_prob(home_team, away_team)) |>
  ungroup() |>
  arrange(desc(winner_prob))

this_week_preds = data.frame(desc = NULL)

for (i in 1:nrow(this_week)) {
  str = paste0(this_week$away_team[i], " @ ", this_week$home_team[i], ": ",
               this_week$pred_winner[i], " def. ", this_week$pred_loser[i], " (", this_week$winner_prob[i], ")")
  
  this_week_preds = rbind(this_week_preds, data.frame(desc = str))
}
```

- `r this_week_preds$desc[1]`
- `r this_week_preds$desc[2]`
- `r this_week_preds$desc[3]`
- `r this_week_preds$desc[4]`
- `r this_week_preds$desc[5]`
- `r this_week_preds$desc[6]`
- `r this_week_preds$desc[7]`
- `r this_week_preds$desc[8]`
- `r this_week_preds$desc[9]`
- `r this_week_preds$desc[10]`
- `r this_week_preds$desc[11]`
- `r this_week_preds$desc[12]`
- `r this_week_preds$desc[13]`
- `r this_week_preds$desc[14]`
<!-- - `r this_week_preds$desc[15]` -->
<!-- - `r this_week_preds$desc[16]` -->



```{r}
# team pts scored each week slugs
end_games

get_week_pts_scored = function(team, wk) {
  x = end_games |> filter((home_team == team | away_team == team) & week == wk)
  if (nrow(x) == 0) return(NA)
  if (x$home_team == team) score = x$home_score else score = x$away_score
  return(score)
}

weekly_scored = data.frame()

for (i in 1:length(all_teams)) {
  for (j in 1:max(end_games$week)) {
    x = data.frame(team = all_teams[i], week = j)
    weekly_scored = rbind(weekly_scored, x)
  }
}

weekly_scored |>
  rowwise() |>
  mutate(week_scored = get_week_pts_scored(team, week)) |>
  ungroup() |>
  na.omit() |>
  mutate(total_scored = sapply(team, get_pts_scored),
         pct = week_scored / total_scored) |>
  ggplot(aes(week, pct)) +
  geom_line(aes(col = team), linewidth = 1.5, show.legend = F) +
  scale_color_manual(values = team_hex) +
  facet_wrap(vars(team), scales = "free_x", nrow = 5) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.5), labels = scales::percent) +
  labs(x = "Week", y = "Percent of Total Points Scored",
       title = "Offensive Scoring Trends")
```

### Team Margins by Half

```{r echo = F}
# points scored and allowed by half
halftime_res = season_pbp |>
  filter(desc == "END QUARTER 2") |>
  select(game_id, home_team, away_team,
         home_half_score = total_home_score, away_half_score = total_away_score)

get_half_pts_scored = function(team) {
  return(halftime_res |>
    filter(home_team == team | away_team == team) |>
    mutate(team_score = ifelse(home_team == team, home_half_score, away_half_score)) |>
    pull(team_score) |>
    sum())
}

get_half_pts_allowed = function(team) {
  return(halftime_res |>
    filter(home_team == team | away_team == team) |>
    mutate(team_score = ifelse(home_team == team, away_half_score, home_half_score)) |>
    pull(team_score) |>
    sum())
}

halves_scored_allowed = data.frame(team = all_teams) |>
  mutate(total_scored = sapply(team, get_pts_scored),
         total_allowed = sapply(team, get_pts_allowed),
         half_scored = sapply(team, get_half_pts_scored),
         half_allowed = sapply(team, get_half_pts_allowed),
         second_half_scored = total_scored - half_scored,
         second_half_allowed = total_allowed - half_allowed)

halves_scored_allowed |>
  mutate(h1_margin = half_scored - half_allowed,
         h2_margin = second_half_scored - second_half_allowed,
         abs_diff = abs(h1_margin - h2_margin)) |>
  select(team, h1_margin, h2_margin, abs_diff) |>
  ggplot(aes(h1_margin, h2_margin)) +
  geom_point(aes(col = team), size = 4, show.legend = F) +
  scale_color_manual(values = team_hex) +
  ggrepel::geom_text_repel(aes(label = team), size = 3.5) +
  geom_hline(aes(yintercept = mean(h2_margin)), linetype = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = mean(h1_margin)), linetype = "dashed", alpha = 0.5) +
  labs(x = "First Half Margin", y = "Second Half Margin",
       title = "Scatterplot of Team Margins by Half") +
  scale_x_continuous(breaks = seq(-100, 100, by = 10)) +
  scale_y_continuous(breaks = seq(-100, 100, by = 10))
```

















